---
title: "Bug Count Analysis"
output:
  html_notebook:
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r}
library(bugcount)
data_dir <- file.path(system.file(package = "bugcount"),"extdata")

```

# Whitefly analysis.
## Read WF count data. 
WF counts per picture. Each picture must have a unique combination of  
experiment + replicate + plant + pic  
experiment description consists of:  
exp_year + exp_cross + exp_propagation + exp_substrate  
```{r}
wf_file <- file.path(data_dir, "wf_consolidated.tab")
wf <- read_wf(file = wf_file, sep = "\t",header=TRUE)
```
## Counts per Plant  

```{r}

wf_count <- plant_count(wf)
summary(wf_count$nymphs)

```

## Internal check analysis.

```{r, warning=FALSE}
wf_check <- wf_count[grepl("internal_check", wf_count$group) &
                       wf_count$clone != "Secundina",]

wf_check <- remove_levels(wf_check)

checks_count_fit <-  count_fit(wf_check)

plot_count_fit(checks_count_fit, main = "Nymphs")
```

Assume that nymphs per plant are distributed as negative binomial

```{r}
by_experiment <- fit_nb_glm(nymphs ~ experiment, wf_check)

plot_fit_nb_glm(wf_check,by_experiment, type = "violin", cld = FALSE)
```
add infestation regime as factor 
high,low: inferred from nyphs ~ experiment GLM "see below"

```{r}
wf_check$infestation <- "low"
wf_check[wf_check$exp_year > 2016,]$infestation <- "high"
wf_check$infestation <- factor(wf_check$infestation, levels =c("low","high"))
```


### Correlation  between experiments

```{r, fig.asp=1}
check_by_exp <- wf_wide(clone ~ experiment, wf_check, 
                        fun = function(x) log10(geometric_mean(x)) )
plot_check_cor(check_by_exp, main = paste("Checks", cor_title))
```



```{r}
# nymphs ~ clone GLM posthoc and common letter difference
fit <- fit_nb_glm(nymphs ~ clone, wf_check)
plot_fit_nb_glm(wf_check,fit,type = "violin", xmax = 10000)
```

```{r, fig.asp=0.5}
fit <- MASS::glm.nb(nymphs ~  infestation * clone, data = wf_check)
posthoc <- lsmeans::cld(lsmeans::lsmeans(fit, ~ infestation * clone,
                                adjust = "tuckey"), type = "response")
# Plot post hoc

### Order the levels for printing
clone_order <- wf_aggregate(wf_check, by_stat = "mean")["clone"]

posthoc$clone <- factor(posthoc$clone, levels = clone_order$clone)

posthoc$infestation = factor(posthoc$infestation,
                   levels=c("low", "high"))

###  Remove spaces in .group  

posthoc$.group=gsub(" ", "", posthoc$.group)

# quartz()
plot_gg_infestationXclone(posthoc)
```


What indicator should we use as resistance phenotype?
Nymph counts are not normal 
Error is not distributed normally.
Nymph counts do not meet linear model assumptions (ANOVA, MapQTL).
Best fit is a negative Binomial Distribution

## Whole population.

### Probability density model fit to actual distribution
```{r, warning=FALSE}
wf_count_fit <-  count_fit(wf_count)

plot_count_fit(wf_count_fit, main = "Nymphs")
```

### Reproducibility 

Do insect counts change per experiment?

nymphs ~ experiment GLM, posthoc and common letter difference 

```{r}
by_experiment <- fit_nb_glm(nymphs ~ experiment, wf_count)

plot_fit_nb_glm(wf_count,by_experiment, type = "violin", cld = FALSE)
```
add infestation regime as factor 
high,low: inferred from nyphs ~ experiment GLM "see below"

```{r}
wf_count$infestation <- "low"
wf_count[wf_count$exp_year > 2016,]$infestation <- "high"
wf_count$infestation <- factor(wf_count$infestation, levels =c("low","high"))
```
### Correlation  between experiments 

```{r, fig.asp=1}

cor_title <- paste("Geometric Means Correlation","Log Scale", sep = "\n")

for (cross in levels(wf_count$exp_cross)) {
  
  wf_by_exp <- wf_wide(clone ~ experiment,
                       wf_count[wf_count$exp_cross == cross,])
  
  plot_wide_cor(wf_by_exp, main = paste(cross,cor_title))
  
}
```



### Nymph counts by Cross 
```{r, fig.asp=1}
exp_levels <- levels(as.factor(wf_count$experiment))
exp_grp <- list(CM8996 = list(1,2,4,5),
                GM8586 = list(3,6,7))
plot_list <- list()
for (cross in levels(wf_count$exp_cross)) {
  for (idx in 1:length(exp_grp[[cross]])) {
  exp_allowed <- exp_levels[exp_grp[[cross]][[idx]]]
  
  wf_by_cross <- wf_count[wf_count$exp_cross == cross &
                            wf_count$experiment %in% exp_allowed,]
  exp_count <- aggregate(experiment ~ clone ,wf_by_cross,
                         FUN = function(x) length(unique(x)))
  selected_clones <- with(exp_count, {
    clone[experiment == length(exp_allowed)]
  })
  
  wf_by_cross <- wf_by_cross[ wf_by_cross$clone %in% selected_clones,]
  wf_by_cross <- (wf_by_cross)
  
  wf_by_cross$nymphs <- wf_by_cross$nymphs+1
  # GLM ************************************************************ #
  # Assuming Negative Binomial distribution
  #
  
  fit_nb <- MASS::glm.nb(nymphs ~  clone, data = wf_by_cross)
  AIC(fit_nb)

  posthoc <- lsmeans::cld(
  lsmeans::lsmeans(
     fit_nb, ~ clone, adjust = "tuckey"),type = "response"
  )
  
  # Plot post hoc
  breaks <- posthoc$clone[!grepl("GM|CM", posthoc$clone, perl = TRUE)]
  
  plot_list[[paste(exp_allowed)]] <- plot_cross_nb_fit(posthoc,
                                                       exp_allowed,
                                                       breaks = breaks)
  
  }
}
ggplot2::theme_set(ggplot2::theme_gray(base_size = 10))
ordered_list <- c(plot_list[1],list(empty = NULL),plot_list[c(3,7,2,5,4,6)])
names(plot_list)
names(ordered_list)
cowplot::plot_grid(plotlist = ordered_list,
          ncol = 2, nrow = 4, align = "v")

```
